Methods of detecting graph similarity.

\begin{itemize}
	\item Widely used in chemistry to predict toxicity and in social network analysis.
	\item 3D-object recognition
	\item conceptual graphs in automatic textual analysis
\end{itemize}

Regularity principle: similar tasks should have similar solutions.
Graph similarity is defined in terms of a similarity function
\[\rho:\Gcal\times \Gcal\ \to \big[0,1\big]\]

\noindent\textbf{Problem statement} \hfill\\
Minimize the functional on the training sample
\[\min_\rho \sum_{i,j} \big(1-\rho(G_i,G_j)\big) 1_{c(G_i)=c(G_j)} + \rho(G_i, G_j) 1_{c(G_i)\neq c(G_j)} \]
where $\big( G_i, c(G_i)\big)_{i=1}^n$ is a class labelled training sample.

Exact methods and approximate methods.

Tree search includes isomorphism

Labelled graph isomorphism $G_1\sim G_2$ iff there exists $\phi:V_1\to V_2$ such that 
\[ \big(\phi(u),\phi(v)\big)\in E_2 \Leftrightarrow (u,v)\in E_1\]
and $L_v = L_{\phi(v)}$ for all $v\in V_1$ and conversely.

Graph isomorphism as a boolean function is a poor similarity function.
Exact isomorphism testing on real world data is not used due to many reasons: \begin{itemize}
	\item Complexity;
	\item Missing data;
	\item Label noise.
\end{itemize}

Relaxation

Subgraph isomorphism based on the maximal isomorphic subgraph of $G_1$ and $G_2$:
\[\rho(G_1,G_2) = 1 - \frac{\lvert\text{mcs}(G_1, G_2)\rvert}{\max(|G_1|,|G_2|)}\]
this is a \textbf{real} metric on the graph space. This problem reduces to maximal clique detection.

Different graph isomorphism algorithms \begin{itemize}
	\item Ulhman's algorithm;
	\item VF2;
	\item Bron-Kerbosch.
\end{itemize}

Ulhman's algorithm works on undirected unlabelled graphs.
It is important to notice that a vertex can be mapped to another vertex only if their degrees match.
The basic idea is to narrow the search space only to the so called candidates.

Graph edit distance is what graph isomorphisms reduce to.

Continuous optimisation. Construct a vector representing a given graph in some space, then run some optimisation procedure and finally reconstruct the graph.

Spectral graph similarity detects the structural similarity.


