Everything you wanted to know about microbiology, but were afraid to ask.

\section{A central dogma} % (fold)
\label{sec:a_central_dogma}
% Watson and Krick -- discovery of the DNA
DNA produces RNA which produces proteins. Molecular biologists are mostly
interested in proteins. Transport proteins, structural proteins.

DNA and RNA store and transfer information. Basically, from the computer
science DNA is just a long string of nucleotides:
\[\ldots \to A\to G\to C\to T\to \ldots \]
Some nuclei are known to contain cyclical DNA sequences.

Two-strains permits error correction and reproduction of the DNA sequence.

RNA is a single strain molecule with the same nucleotides, except for
$U$ is substituted for $T$.

Two kinds of RNA: \begin{itemize}
	\item regulatory RNA;
	\item matrix, transport RNA.
\end{itemize}

Every RNA is a copy of some DNA subsequence. DNA is like a spool storage,
whereas RNA is cache for protein synthesis.

There are 20 amino-acids, which are encoded by triples of nucleotides.

RNA can self organize in line with the complementary nucleotides:
\[\ldots-UUACGU-\ldots -ACGUAA-\ldots\]
can fold into a spindle along $A-U$ and $G-C$ complements.

Bacteria have special proteins: DNA polymerase which can recognize certain
DNA subsequences.
\begin{itemize}
	\item DNA promoter, which identifies specific subsequences;
	\item DNA terminator, which acts like a stop-signal for the polymerase.
\end{itemize}

The process of synthesising RNA from DNA is \textbf{transcription}.

Ribosome consists of structural RNA with a lot of different proteins. Ribosome
reads a triple of nucleotides, and then attaches an appropriate aminoacid to 
the protein. There are special coding triples signifying synthesis stopping.
This is called \textbf{translation}.

% Transport-RNA 

% section a_central_dogma (end)
